---
title: "Perception Non-native Accented Speech"
author: ""
---

```{r, echo=FALSE}
library(dplyr)
library(effects)
library(lattice)
library(lme4)
library(optimx) # use different optimizer (bobyqa) than the lme4 default one
library(phonR)
library(scales)
```
Load data 

```{r}
formanti <- read.csv2("formanti.csv", na.strings = NA, stringsAsFactors = TRUE)
formantie <- read.csv2("formantie.csv", na.strings = NA, stringsAsFactors = TRUE)
```

Figure with formants per vowel

```{r, fig.height=4,fig.width=10}
par(mfrow=c(1,2))
with(formanti, plotVowels(F1, F2, group = Condition, var.col.by = Condition, 
                          plot.tokens = FALSE, plot.means = TRUE, pretty=TRUE))
text(1900, 410, "canonical ɪ")
with(formantie, plotVowels(F1, F2, group = Condition, var.col.by = Condition, 
                           plot.tokens = FALSE, plot.means = TRUE, pretty=TRUE))
text(2400, 300, "canonical i")
par(mfrow=c(1,1))
```
# Lexical Decision Task

Load data

```{r, echo=FALSE}
ldt <- read.csv2("LexDecTask.csv", na.strings = NA, stringsAsFactors = TRUE)
```

```{r, echo=FALSE}
summary(ldt)
```

Select target words

```{r, echo=FALSE}
ldt <- ldt %>% filter(Item_Type == "target_word_i" | Item_Type == "target_word_ie")
ldt <- droplevels(ldt)
```

Rename variables and levels

```{r, echo=FALSE}
ldt<-rename(ldt,
  "TargetWord" =	Item_Type)
levels(ldt$TargetWord) <- c("ɪ", "i")
```

```{r, echo=FALSE}
levels(ldt$Condition) <- c("ɪ ambiguous", "i ambiguous")
```

```{r, echo=FALSE}
summary(ldt)
```

3.2.1 Word endorsement: logistic regression

```{r}
mm1 <- glmer(Part_Response ~ TargetWord*Condition + (1|Participant) + (1|Item), 
             control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
             data = ldt, 
             family=binomial)
mm2 <- glmer(Part_Response ~ TargetWord + Condition + (1|Participant) + (1|Item),
             control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), 
             data = ldt, 
             family=binomial)
anova(mm1, mm2)
```

```{r, echo=FALSE}
summary(mm1)
```

Effect plot of Interaction

```{r, echo=FALSE}
plot(allEffects(mm1), ylim=c(0,1), main="", rescale.axis=FALSE, ylab="Prob(Decision = Word)")
```

```{r}
png("Figure_2.png", width=15, height=8, units="cm", res=450)
plot(allEffects(mm1), ylim=c(0,1), main="", rescale.axis=FALSE, 
     xlab="Target word",
     ylab="Prob(Decision = Word)")
dev.off()
```

Plot of random effect Item

```{r, echo=FALSE}
dotplot(ranef(mm1, which = "Item", condVar = TRUE))
```

# Response time

Figure 3: boxplots comparing RT_ms by Condition and TargetWord

```{r}
p <- ldt %>% ggplot(aes(x=TargetWord, y=RT_ms, colour=TargetWord)) +
  theme_bw() +
  geom_point(position=position_jitterdodge(jitter.width = 0.50), alpha=0.3, size=1) + 
  stat_summary(fun=mean, geom="point", shape=23, size=3, colour="darkblue", fill="white") +
  scale_colour_brewer(palette = "Set1", guide = FALSE) +
  facet_wrap(vars(Condition)) + #, labeller = Condition_names) + 
  xlab("Target word") + 
  ylab("Response time (ms)") +
  scale_x_discrete(labels=c("ɪ word", "i word")) + 
  theme(strip.text = element_text(size=rel(1.3)))
  p
```

```{r}
ggsave("Figure_3.png", width = 16, height = 9, units = "cm")
```

Regression model

```{r, echo=FALSE}
m1 <- lmer(sqrt(RT_ms) ~ TargetWord*Condition + (1|Participant) + (1|Item), ldt)
m2 <- lmer(sqrt(RT_ms) ~  TargetWord + Condition + (1|Participant) + (1|Item), ldt)
anova(m1, m2)
```

Interaction is significant

Model quality

fitted vs resid

```{r, echo=FALSE}
plot(m1,type=c("p","smooth"))
```

Scale-location

```{r, echo=FALSE}
plot(m1,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth"))
```

qqplot residuals

```{r, echo=FALSE}
qqmath(m1)
```

Effect plot of Interaction

```{r, echo=FALSE}
png("Figure_4.png", width=15, height=8, units="cm", res=450)
plot(allEffects(m1), main="", xlab="Target word", ylab="Response time (sqrt(ms))")
dev.off()
```

# 3.2.2	Response time 

Load data

```{r, echo=FALSE}
ldt <- read.csv2("LexDecTask.csv", na.strings = NA, stringsAsFactors = TRUE)
```

Select target words

```{r, echo=FALSE}
ldt <- ldt %>% filter(Item_Type == "target_word_i" | Item_Type == "target_word_ie")
ldt <- droplevels(ldt)
```

Select correct responses

```{r, echo=FALSE}
ldt <- ldt %>% filter(Response == "correct")
ldt <- droplevels(ldt)
```

Rename variables and levels

```{r, echo=FALSE}
ldt<-rename(ldt,
  "TargetWord" =	Item_Type)
levels(ldt$TargetWord) <- c("i", "ie")
```

```{r, echo=FALSE}
levels(ldt$Condition) <- c("i ambiguous", "ie ambiguous")
```

Regression model

```{r, echo=FALSE}
m1 <- lmer(sqrt(RT_ms) ~ TargetWord*Condition + (1|Participant) + (1|Item), ldt)
m2 <- lmer(sqrt(RT_ms) ~  TargetWord + Condition + (1|Participant) + (1|Item), ldt)
anova(m1, m2)
```

Model quality

Fitted vs resid

```{r, echo=FALSE}
plot(m1,type=c("p","smooth"))
```

Scale-location

```{r, echo=FALSE}
plot(m1,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth"))
```

QQ-plot of residuals

```{r, echo=FALSE}
qqmath(m1)
```

Effect plot of Interaction

```{r, echo=FALSE}
plot(allEffects(m1))
```
# Phonological categorisation Task

Load data

```{r, echo=FALSE}
pct <- read.csv2("PhonCatTask.csv", na.strings = NA, stringsAsFactors = TRUE)
```

Change Step and Iteration to factor

```{r, echo=FALSE}
pct$Step <- as.factor(pct$Step)
pct$Iteration <- as.factor(pct$Iteration)
```

Summarize

```{r, echo=FALSE}
summary(pct)
```

Vowel vs. Response

```{r, echo=FALSE}
tbl <- table(pct$Condition, pct$Response)
tbl
```

Figure 5

```{r}
pp <- pct %>% group_by(Step, VoiceCond) %>% count(Response) %>% 
  mutate(freq = n / sum(n), label = percent(freq %>% round(2))) %>%
  filter(Response  == "ie") %>%
  ungroup()
pp
```

Change level names

```{r, echo=FALSE}
levels(pp$VoiceCond) <- c("female /ɪ/", "female /i/", "male /ɪ/", "male /i/")
```

```{r}
p <- ggplot(pp, aes(x = Step, y=freq, colour=VoiceCond, group=VoiceCond)) +
  geom_line(size=1.1) +
  geom_point(size=3, shape=21, fill="white") +
  theme_bw() +
  scale_colour_brewer(palette="Set1", guide = FALSE) +
  xlab("Step") +
  ylab("Proportion /i/-responses") +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  annotate("text", x = 4.05, y = 0.44342, hjust=0, vjust=0, label = "male /i/") + # add names to lines
  annotate("text", x = 4.05, y = 0.38, hjust=0, vjust=0, label = "male /ɪ/") + # add names to lines
  annotate("text", x = 4.05, y = 0.15345, hjust=0, vjust=0, label = "female /i/") + # add names to lines
  annotate("text", x = 4.05, y = 0.1375, hjust=0, vjust=1, label = "female /ɪ/") # add names to lines
p  
```
```{r}
ggsave("Figure_5.png", width = 16, height = 9, units = "cm", dpi = 450)
```

Mixed effects log regr model

Forward difference coding

https://stats.idre.ucla.edu/r/library/r-library-contrast-coding-systems-for-categorical-variables/#forward

```{r, echo=FALSE}
my.forward.diff = matrix(c(3/4, -1/4, -1/4, -1/4, 1/2, 1/2, -1/2, -1/2, 1/
	4, 1/4, 1/4, -3/4), ncol = 3)
my.forward.diff
```

```{r, echo=FALSE}
contrasts(pct$Step) <- my.forward.diff
```

```{r}
str(pct$Response)
```
ie is outcome level

```{r, echo=FALSE}
mm1 <- glmer(Response ~ Voice*Condition + Step + (1|Participant) + (1|Item), data = pct, family=binomial)
mm2 <- glmer(Response ~ Voice + Condition + Step + (1|Participant) + (1|Item), data = pct, family=binomial)
anova(mm2, mm1)
```

The interaction is not significant.

```{r, echo=FALSE}
summary(mm2)
```

- Voice: male voice is has a positive effect on the /ie/ categorization 
- each adjacent step is associated with a significantly larger probability of /ie/. Largest effect between step 3 and 4. 

```{r, echo=FALSE}
png("Figure_6.png", width=15, height=8, units="cm", res=400)
plot(allEffects(mm2)[c(1,3)], ylim=c(0,1), main="", rescale.axis=FALSE, ylab="Prob(Cat = /i/)")
dev.off()
```

```{r, echo=FALSE}
png("Figure_7.1.png", width=20, height=8, units="cm", res=400)
dotplot(ranef(mm2, which = "Item", condVar = TRUE))
dev.off()
```

```{r}
png("Figure_7.2.png", width=20, height=15, units="cm", res=400)
dotplot(ranef(mm2, which = "Participant", condVar = TRUE))
dev.off()
```